---
title: "p8105_hw5_ns3923"
output: github_document
---

**Set up necessary library**
```{r setup, message=FALSE}
library(tidyverse)
```

## Problem 1

**Goal:** To estimate the probability that at least two people share a birthday for group sizes 2–50 using 10,000 simulations each.

**1) Define a function that simulates random birthdays for a group of size n and returns TRUE if any birthdays are shared.**

```{r P1_1}
set.seed(123)

dup_bday = function(n, days = 365) {
  birthdays = sample.int(days, size = n, replace = TRUE)
  any(duplicated(birthdays))
}

#Check
test_results =
  tibble(
    group_size = c(2, 10, 15, 60),
    shared_bday = map_lgl(group_size, dup_bday)
  )
test_results
```
From the checking table above, smaller group sizes such as 2 and 10 people are less likely to have a shared birthday, as shown by `shared_bday = FALSE`. In contrast, larger groups, especially with 60 people, have a higher chance of at least one shared birthday (`shared_bday = TRUE`). This makes sense because the more people in a group, the greater the chance that two will share the same birthday.

**2) Repeat the simulation 10,000 times for each group size between 2 and 50, and calculate the probability of a shared birthday**
```{r P1_2}
Repeat_group =
  expand_grid(group_size = 2:50, iter = 1:10000) |>
  mutate(dup = map_lgl(group_size, dup_bday)) |>
  group_by(group_size) |>
  summarize(prob_duplicate = mean(dup), .groups = "drop")
```

**3) Make a plot showing the probability as a function of group size**

```{r P1_3}
Repeat_group |>
  ggplot(aes(x = group_size, y = prob_duplicate)) +
  geom_line() +
  geom_point(size = 2, color = "skyblue") +
  labs(
    title = "Problem 1: Probability of at least 2 people shared birthday",
    x = "Group Size (n)",
    y = "Probability"
  ) +
  theme_minimal()
```

**Explain:** The trend is upward as the group size increases, meaning that larger groups have a higher probability of two people sharing a birthday. The probability rises and approaches 1 as the group becomes larger. Around 23 people, the chance of a shared birthday is about 50%.

